---
title: "W4 Project2 NOAA storm database"
author: "Khalid"
date: "7/12/2020"
output: html_document
---

### Your data analysis must address the following questions:

    1. Across the United States, which types of events (as indicated in the color red  variable) are most 
       harmful with respect to population health?
    2. Across the United States, which types of events have the greatest economic consequences?
    
     Consider writing your report as if it were to be read by a   
     government or municipal manager who might be responsible for 
     preparing for severe weather events and will need to 
     prioritize resources for different types of events. However, 
     there is no need to make any specific recommendations in your 
     report.
### Load the libraries

```{r}
library(dplyr)
library("data.table")
library("ggplot2")
```


### Download and extract the data into a dataframe.Then convert to a data.table

```{r}
fileUrl <- "https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2FStormData.csv.bz2"
download.file(fileUrl, destfile = paste0(getwd(), '/Storm.csv.bz2'), method = 'curl', quiet = T)
storm_DB <- read.csv('Storm.csv.bz2', stringsAsFactors = F)

```
### Converting data.frame to data.table
```{r}
storm_DB <- as.data.table(storm_DB)

```
### Examining DATA
```{r}
names(storm_DB)
```
### OR 
```{r}
colnames(storm_DB)
```


### Subset the dataset on the parameters that we need and remove unneeded columns.

    - EVTYPE: event type  
    - FATALITIES: number of fatalities  
    - INJURIES: number of injuries  
    - PROPDMG: property damage (dollars)  
    - PROPDMGEXP: magnitude of property damage (K = thousands, M =       millions, B = billions)  
    - CROPDMG: crop damage (dollars)  
    - CROPDMGEXP: magnitude of crop damage (H = hundreds, K =    
    thousands, M = millions, B = billions)
### Select columns
```{r}
col_select <- c('EVTYPE', 'FATALITIES', 'INJURIES', 'PROPDMG', 'PROPDMGEXP', 'CROPDMG', 'CROPDMGEXP')

storm1 <- storm_DB[, ..col_select]
summary(storm1)
```


### Find types of events that are most harmful with respect to population health
```{r}

stormDB <- storm1[, 1:3] %>% group_by(EVTYPE) %>% summarise_all(sum)

```

```{r}
summary(stormDB)
```
### Top 10 events caused death
```{r}
top10Death <- stormDB[order(stormDB$FATALITIES, decreasing = T), ]
head(top10Death)
```

### Top 10 events caused injuries
```{r}
top10inj <- stormDB[order(stormDB$INJURIES, decreasing = T), ]
head(top10inj)

```
### Top 10 events caused both death & injuries
```{r}
stormDB$total <- rowSums(stormDB[, 2:3])
top10both <- stormDB[order(stormDB$total, decreasing = T), ]
top10both[1:10,]
```

```{r}
head(top10both)
```


### numbers, characters are all represent factors

```{r}
unique(storm1$PROPDMGEXP)
```



### a function to transform 
```{r}
getv <- function(EXP_Type) {
  if (EXP_Type %in% c('h', 'H')) {
    return(2)
          
  } else if (EXP_Type %in% c('k', 'K')) {
    return(3)
          
  } else if (EXP_Type %in% c('m', 'M')) {
    return(6)
          
  } else if (EXP_Type %in% c('b', 'B')) {
    return(9)
          
  } else if (suppressWarnings(!is.na(as.numeric(EXP_Type)))) {
    
    return(as.numeric(EXP_Type))
          
  } else {
    return(0)
  }
}
c(10**getv('h'), 10**getv(4), 10**getv('B'), 10**getv('?'))
```
### Make a table & Put Result into a Table
```{r}
newST <- storm1[, c(1, 4:7)] %>%
  rowwise() %>%
  mutate(PROP = PROPDMG*10**getv(PROPDMGEXP), 
         CROP = CROPDMG*10**getv(CROPDMGEXP))

```


```{r}
head(newST)
```

```{r}
newST_Sum <- newST[, c(1, 6, 7)] %>%
  group_by(EVTYPE) %>%
  summarise_all(sum)
```

```{r}
summary(newST_Sum)
```

```{r}

```

### top 10 events that causes the most damages
```{r}
topP <- newST_Sum[order(newST_Sum$PROP, decreasing = T), ]
```


### top 10 events that causes the most crop damages
```{r}
topCrop <- newST_Sum[order(newST_Sum$CROP, decreasing = T), ]
```

### combine the top 10 most damages & crop damages
```{r}
newST_Sum$total <- rowSums(newST_Sum[, 2:3])
combine_top <- newST_Sum[order(newST_Sum$total, decreasing = T), ]
```

```{r}
top10Death[1:10,]

```

```{r}
top10inj[1:10,]
```


```{r}
top10both[1:10,]
```


```{r}
ggplot(data = top10both[1:10,], aes(x = reorder(EVTYPE, total), y = total)) +
  geom_bar(stat = 'Identity') +
  coord_flip() +
  xlab('Event Type') +
  ylab('Total Injuries and Fatalities') +
  ggtitle('Top 10 weather events that causes population health hazards') +
  theme_classic()
```


```{r}
ggplot(data = combine_top[1:10,], aes(x = reorder(EVTYPE, total), y = total)) + 
  geom_bar(stat = 'Identity') +
  coord_flip() +
  scale_y_continuous(trans = 'log10') +
  xlab('Event   Type') +
  ylab('Total property and crop damages (log10)') +
  ggtitle('Top 10 Weather events that causes Economic Hazards') +
  theme_classic()
```
```{r}
summary(combine_top)
```





