---
title: "Regression Models Course Project"
author: "Khalid"
date: "7/17/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### You work for Motor Trend, a magazine about the automobile industry. Looking at a data set of a collection of cars, they are interested in exploring the relationship between a set of variables and miles per gallon (MPG) (outcome). They are particularly interested in the following two questions:

    “Is an automatic or manual transmission better for MPG”
    "Quantify the MPG difference between automatic and manual    
     transmissions"




```{r}
data("mtcars")
library(ggplot2)
library(dplyr)
```

### manual transmission has higher mean of miles per gallon
```{r}
auto_mpg <- mtcars %>% select(mpg, am) %>% mutate(am = as.factor(am))
ggplot(data = auto_mpg, aes(x = am, y = mpg)) +
        geom_boxplot() + geom_point() +
        xlab('Transmission (0 = automatic, 1 = manual)') + 
        ylab('Miles/(US) gallon')
```

### p-value less than 5% in the full model
```{r}
D <- mtcars %>% mutate(cyl = as.factor(cyl), vs = as.factor(vs), am = as.factor(am), gear = as.factor(gear), carb = as.factor(carb))
fit_All <- lm(mpg ~ ., data = D)
summary(fit_All)$coef[, 4]
```

# Still no coefficients with a significant p-value after removing the cyl variable
```{r}
which.max(summary(fit_All)$coef[, 4]) 
fitD <- D %>% select(-cyl); fitRm <- lm(mpg ~ ., data = fitD); summary(fitRm)$coef[, 4]
```

```{r}
which.max(summary(fitRm)$coef[, 4]) #the carb variable
```

```{r}
summary(fitRm)$coef[, 4]; which.max(summary(fitRm)$coef[, 4]) #the gear variable
```

### the vs variable
```{r}
summary(fitRm)$coef[, 4]; which.max(summary(fitRm)$coef[, 4]) 
fitD <- fitD %>% select(-vs); fitRm <- lm(mpg ~ ., data = fitD)
```

### the drat variable
```{r}
summary(fitRm)$coef[, 4]; which.max(summary(fitRm)$coef[, 4]) 
fitD <- fitD %>% select(-drat); fitRm <- lm(mpg ~ ., data = fitD)

```
### the disp variable
```{r}
summary(fitRm)$coef[, 4]; which.max(summary(fitRm)$coef[, 4]) 
fitD <- fitD %>% select(-disp); fitRm <- lm(mpg ~ ., data = fitD)
```

### the hp variable
```{r}
summary(fitRm)$coef[, 4]; which.max(summary(fitRm)$coef[, 4]) 
fitD <- fitD %>% select(-hp); fitRm <- lm(mpg ~ ., data = fitD)
```


```{r}
summary(fitRm)$coef[, 4]
```

### Clearly the p-values less than 0.05. three coefficients, wt, qseq, and Transmission
```{r}
summary(fitRm)
```

```{r}
par(mfrow = c(2, 2))
plot(fitRm)
```

# The Conclusion:
       
    the manual one will have an average of 2.9358 higher miles/gallon than the automatic car

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

