---
title: "Swiss Fertility Data Exploration"
author: "Khalid"
date: "8/11/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction
This is the final course porject of Coursera Data Science Specialization Course: Developing Data Science Tools.

1. The Swiss Fertility Data (swiss data) can be accessed with **data(swiss)** in R
2. My GitHub repository contais the codes for Shiny App (**server.R** and **ui.R**) and you could find it [here](https://github.com/khalidm1/datasciencecoursera/tree/master/C%209%20Developing%20Data%20Products)
3. The Shiny app can be accessed [here](https://khlaidm1.shinyapps.io/swiss/), which contains
  

```{r}
data(swiss)
summary(swiss)
```

## Histogram
```{r}
inputVar <- 'Fertility'; inputBin <- 10; histVal <- swiss[, inputVar]
hist(histVal, breaks = seq(min(histVal), max(histVal), length.out = inputBin+1),
     xlab = inputVar, main = paste('Distribution of', inputVar),
     col = 'darkgray', border = 'white')
```

## Scatter Plot
```{r}
library(ggplot2)
scatX <- 'Fertility'; scatY <- 'Education'; scatC <- 'Examination'
ggplot(data = swiss, aes(x = Fertility, y = Education, color = Examination)) + 
      geom_point() + xlab(scatX) + ylab(scatY) + labs(colour = scatC) +
      ggtitle(paste('Scatter plot of', scatX, 'vs', scatY)) +
      theme(plot.title = element_text(hjust = 0.5))
```
